Selecting and ranking interesting pairs of cyclic temporal granularities

### 2021-01-20 paper/main.Rmd
changing paper structure

# The distance measure

We are interested in assessing structure in probability distributions of the measured variable across bivariate cyclic granularities. We propose a measure called Median Maximum Pairwise Distances (wpd) to evaluate structure in such a design.

@@

## Methodology

### Characterising distribution

@@

### Distance between distributions

@@

### Notations and Definitions

Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets. Let the pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$ b   e denoted as $d_{(jk, j'k')} = JSD(a_{j}b_{k}, a_{j'}b_{k'})$. Pairwise distances could be within-facets or between-facets. Figure \ref{fig:distance-explain} illustrates how the within-facet or between-facet distances are defined. Pairwise distances are within-facets ($d_{w}$) when $b_{k} = b_{k'}$, that is, between pairs of the form $(a_{j}b_{k}, a_{j'}b_{k})$ as shown in panel (3) of Figure \ref{fig:distance-explain}. If categories are ordered (like all temporal cyclic granularities), then only distances between pairs where $a_{j'} = (a_{j+1})$ are considered (panel (4)). Pairwise distances are between-facets ($d_{b}$) when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$.

@@

<!-- <!-- 5. Compute distribution of maximum distance ($M$) by shuffling the data $200$ times and finishing Steps1-4 in each case. Scale $M$ using the mean and sd of the distribution.  --> -->

@@

@@

>>>>

# The proposed distance measure

We are interested in assessing structure in probability distributions of the measured variable across bivariate cyclic granularities. We propose a measure called Weighted Maximum Pairwise Distances (wpd) to evaluate structure in such a design.

@@

## Characterising distributions

@@

## Distance between distributions

@@

## Definition of the proposed distance measure

Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets. Let the pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$ be denoted as $d_{(jk, j'k')} = JSD(a_{j}b_{k}, a_{j'}b_{k'})$. Pairwise distances could be within-facets or between-facets. Figure \ref{fig:distance-explain} illustrates how the within-facet or between-facet distances are defined. Pairwise distances are within-facets ($d_{w}$) when $b_{k} = b_{k'}$, that is, between pairs of the form $(a_{j}b_{k}, a_{j'}b_{k})$ as shown in panel (3) of Figure \ref{fig:distance-explain}. If categories are ordered (like all temporal cyclic granularities), then only distances between pairs where $a_{j'} = (a_{j+1})$ are considered (panel (4)). Pairwise distances are between-facets ($d_{b}$) when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$.

@@

<!-- <!-- 5. Compute distribution of maximum distance ($M$) by shuffling the data $200$ times and finishing Steps1-4 in each case. Scale $M$ using the mean and sd of the distribution.  -->

@@

The distribution of wpd is different for different levels of facets and x-axis levels. This is because the statistic maximum which is used to define wpd is affected by the number of categories. The measure would have higher values if $C_i$ or $C_j$ has higher levels. However, we would ideally want a higher value of the measure only if there is significant difference between distributions across facet or x-axis categories, and not because the number of categories are higher. Therefore, in order to compare wpd across different combinations of facet and x-axis levels, we need to eliminate the impact of different levels of the facets and x-axis first and get a normalized measure. Henceforth we call the measure discussed as $wpd_{raw}$ and the normalised measure as $wpd_{norm}$. The measure $wpd_{norm}$ could potentially lead to comparison of the measure across different panels and also identifying only the interesting panels from a dataset.

@@

Most of the behavior of the measure wpd was studied via simulation. The simulations explore how wpd performs under various designs and parameters and its limitations. To study the behavior of wpd, simulations were carried out for four different designs and the following factors that could potentially have an impact on the values of wpd:

<!-- Numerous simulations were carried out in a wide range of scenarios. -->

<!-- Same scenarios were used, when possible with normalized wpd, keeping all other simulation variables constant. -->

- $nx$ (number of levels of x-axis)
- $nfacet$ (number of levels of facets)
- $\lambda$ (tuning parameter)
- $\omega$ (increment in each panel design)
- $dist$ (normal/non-normal distributions with different location and scale)
- $n$ (sample size for each combination of categories)
- $nsim$ (number of simulations)
- $nperm$ (number of permutations of data)  
- $designs$  
$D_{null}$  (No difference in distribution)  
$D_{var_f}$ (Difference in distribution only across facets)  
$D_{var_x}$ (Difference in distribution only across x-axis)  
$D_{var_{all}}$ (Difference in distribution in both facets and x-axis)

### Environment

R version 4.0.1 (2020-06-06) is used with platform: x86_64-apple-darwin17.0 (64-bit) running under: macOS Mojave 10.14.6 and MonaRCH, which is a next-generation HPC/HTC Cluster, designed from the ground up to address the computing needs of the Monash HPC community. The nodes and the storage used are as follows:

### Results

This section reports the results of a simulation study that was carried out to evaluate the
behavior of our distance measure across all potential factors. The results are reported in two parts: for the a) raw measure and then b) normalised one to show how the loopholes for the raw measures were removed using the normalized ones.

First, the behavior of $wpd_{raw}$ and $wpd_{norm}$ is explored in designs where there is no difference in distribution between x and facet categories. We have considered different initial distributions to study the impact of initial distribution under the null setup. Using two types of distributions, viz. normal and gamma (non-normal), we generated observations for each combination of $nx$ and $nfacet$ from the following sets: $nx = \{2, 3, 5, 7, 14, 20, 31, 50\}$ and $nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$ to cover a wide range of levels from very low to moderately high. Each combination is being referred to as a _panel_. That is, data is being generated for each of the panels $\{nx = 2, nfacet = 2\}, \{nx = 2, nfacet = 3\}, \{nx = 2, nfacet = 5\},  \dots, \{nx = 50, nfacet = 31\}, \{nx = 50, nfacet = 50\}$. For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories. That is, if we consider the panel  $\{nx = 2, nfacet = 2\}$, $500$ observations are generated for each of the combination of categories from the panel, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$. The values of $\lambda$ is set to $0.67$, since we want to up-weigh the within-facet distances and that of $\omega$ is set to $0$, since there is no significant differences between distributions in the null case. Observations were generated for each type of distribution changing the shape and scale to study the effect of shape, scale and type of distribution on wpd. The set of distributions considered for this purpose is ${N(0,1), N(5, 1), N(0,5), \Gamma(0.5, 1), \Gamma(2, 1)}$. Each of the scenario is run $nsim = 200$ times to see the distribution of wpd values for each scenario.

Secondly, the behavior of raw and normalized wpd is explored in designs where there is in fact difference in distribution between facet categories ($D_{var_f}$) or across x-categories ($D_{var_x}$) or both ($D_{var_{all}}$). Using $\omega =  \{1, 2, \dots, 10\}$ and $\lambda = seq(from  = 0.1, to = 0.9, by = 0.05)$, observations are drawn from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet =  \{2, 3, 5, 7, 14, 20, 31, 50\}$. $ntimes = 500$ is assumed for this setup as well. Furthermore, to generate different distributions across different combination of facet and x levels, the following method is deployed - suppose the distribution of the combination of first levels of x and facet category is $N(\mu,\sigma)$ and $\mu_{jk}$ denotes the mean of the combination $(a_jb_k)$, then $\mu_{j.} = \mu + j\omega$ (for design $D_{var_x}$) and $\mu_{.k} = \mu + k\omega$ (for design $D_{var_f}$).

The tabulated values and graphical representations of the simulation results are provided in Appendix. The learning from the simulations are as follows:

## Properties
